\documentclass{article}
\usepackage{url}
\usepackage{amsmath,bm}%
\usepackage{amsfonts}%
\pagestyle{empty}
\setlength{\textwidth}{7in}
\setlength{\oddsidemargin}{-.5in}
\setlength{\evensidemargin}{-.5in}
\setlength{\topmargin}{-.75in}
\setlength{\textheight}{9.25in}

\newcommand{\beaa}{\begin{eqnarray*}}
\newcommand{\eeaa}{\end{eqnarray*}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\def\E{\mathop{\rm E\,}\nolimits}
\def\Var{\mathop{\rm Var\,}\nolimits}
\def\Cov{\mathop{\rm Cov\,}\nolimits}
\def\Corr{\mathop{\rm Corr\,}\nolimits}
\def\logit{\mathop{\rm logit\,}\nolimits}
\newcommand{\eid}{{\stackrel{\cal{D}}{=}}}
\newcommand{\cip}{{\stackrel{{P}}{\to}}}
\def\bR{\mathbb{R}}     % real line

\begin{document}

\begin{center}
{\bf STAT 515}

{\bf Homework \#6, due Friday, Mar. 2 at 2:30pm}

{\bf This homework may be submitted electronically to ANGEL, though this is not
required. I strongly encourage the use of \LaTeX\ in any case.}

\end{center}

{\it Please make every assignment easier to grade by neatly organizing your
writeup and clearly labeling your final answers when appropriate. Try using
\LaTeX!}


\begin{enumerate}

% From Murali's HW#5
\item Let $N(t)$ be a non-homogeneous Poisson process with rate function
$\lambda(t)$.

  \begin{enumerate}

  \item Fix some $t>0$. Derive the conditional distribution of the arrival times
  $S_1, \ldots, S_{N(t)}$, conditional on $N(t)=n$.

  \item As you did in Exercise 2(a) of Homework \#5, use part (a) to suggest an
  algorithm for simulating a non-homogeneous Poisson process on $(0, t]$.

  \item Use your algorithm to simulate 10,000 realizations of a non-homogeneous
  Poisson process on $(0, 2]$ where $\lambda(t) = 24t^2$.

  \item Using the result of part (c), produce a histogram of the numbers of
  events in the interval $(0, 1]$. Add to your histogram the true theoretical
  values, and explain how you found them.

  \item Using the result of part (c), produce a histogram of the times until the
  first event. Add to your histogram the true theoretical density, and explain
  how you found it. What is the name of the true theoretical distribution?

  \end{enumerate}  

% Ex 5.52
\item Teams 1 and 2 are playing a game. The teams score points according to
independent Poisson processes with rates $\lambda_1$ and $\lambda_2$,
respectively. The game ends when one team has scored exactly $k$ points more
than the other team. What is the probability that team 1 wins?

% Ex 5.28
\item Consider $n$ components with independent lifetimes, where component $i$
has exponentially distributed lifetime with rate $\lambda_i$. Suppose that all
components are initially in use and remain so until they fail.

  \begin{enumerate}

  \item Find the probability that component 1 is the second component to fail.

  \item Find the expected time of the second failure.

  \end{enumerate}

\end{enumerate}

\end{document}

