\input{lecture}

\begin{document}


\def \thedate{Jan.~20}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
Read Sections 4.2 and 4.3 (both editions) before Monday's class.
\item
I encourage you to:
\begin{itemize}
\item Read ahead!
\item Browse Section 3.6 (both editions)
\end{itemize}
\item I have made an addition to the syllabus.
\end{itemize}
\end{frame}


\startframe{3.4 Computing Expectations by Conditioning}
\begin{itemize}
\item Perform repeated independent Bernoulli$(p)$ experiments.
\item Let $N_k$ be the trial number where $k$ consecutive successes are first observed.
\item Recall: $E(N_k) = E \left[ E(N_k \mid N_{k-1}) \right]$
\end{itemize}
\end{frame}

\startframe{3.5 Computing Probabilities by Conditioning}
\begin{itemize}
\item 
Two players alternate flipping a coin that comes up heads with probability $p$.
\item
The first one to get heads is the winner.
\item 
What is the probability that the first player is the winner?
\end{itemize}
\end{frame}

\startframe{4.1 Introduction to Markov Chains}

In chapter 4, ``time'' $t$ is discrete:  $t=0, 1, 2, \ldots$

\vspace{2ex}
 ``Markov'' means:  
Distribution of $X_t$ depends {\em only} on $X_{t-1}$.

{\em (Not on $t$ nor on $X_{t-2}, X_{t-3}, \ldots$.)}

\end{frame}


\startframe{4.1 Introduction to Markov Chains}
Consider the simple weather example with states ``rain'' and ``no rain'':
\[
P=
\begin{bmatrix}
\alpha & 1-\alpha \\
\beta & 1-\beta
\end{bmatrix}
=
\begin{bmatrix}
0.7 & 0.3 \\
0.4 & 0.6
\end{bmatrix}
\]
\end{frame}

\startframe{4.1 Introduction to Markov Chains}
More complicated rain example with four states:

(0) rain, rain; (1) rain, no rain; (2) no rain, rain; (3) no rain, no rain

\[
P=
\begin{bmatrix}
0.7 & 0 & 0.3 & 0 \\
0.5 & 0 & 0.5 & 0 \\
0 & 0.4 & 0 & 0.6 \\
0 & 0.2 & 0 & 0.8
\end{bmatrix}
\]
\end{frame}

\startframe{4.1 Introduction to Markov Chains}
\question{If $P$ is the transition matrix for a finite-state Markov chain 
$X_0, X_1, X_2, \ldots$, and if the $X_i$ are i.i.d., then:}
{Each (horizontal) row of $P$ is the same.}
{Each (vertical) column of $P$ is the same.}
{Each column of $P$ sums to 1.}
{None of A, B, or C.}
\end{frame}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


