\input{lecture}

\begin{document}


\def \thedate{Feb.~24}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
For Monday, read Section 6.3.
\item 
Midterm exam:  7:00pm on Wednesday, Feb.~29 in 105 Willard.
(NB:  Change of room!)
\item 
HW \#6 will be due next Friday (Mar.~2) at 2:30.
\end{itemize}
\end{frame}


\startframe{Midterm practice}
\question{Let $X_1$ and $X_2$ be independent exponential random variables with
rates $\lambda_1$ and $\lambda_2$, and let $X_{(1)}$, $X_{(2)}$ be the
order statistics.  What is $\Cov (X_{(1)}, X_{(2)})$?}
{$(\lambda_1+\lambda_2)^2$}
{$\lambda_1\lambda_2$}
{$1/(\lambda_1+\lambda_2)^2$}
{$1/(\lambda_1\lambda_2)$}
\end{frame}

\startframe{6.2 Continuous-Time Markov chains}
Consider a continuous-time, discrete-space MC, $\{X(t) : t\in[0,\infty)\}$:
\begin{itemize}
\item The chain stays in each state for a random (real-valued) time.
\item When the chain switches, it picks from the possible states according to
some discrete-time Markov chain.
\item Markov property:  
Conditional on the present, what happens next is independent of the past.
\item Homogeneity:  The chain's behavior at time $t$ does not
depend on $t$.
\end{itemize}
\end{frame}

\startframe{6.2 Continuous-Time Markov chains}
Let $T_i$ be the time when the MC leaves state $i$,
given that $X(0)=i$.
\begin{itemize}
\item How is $T_i$ distributed?
\item Claim:  $T_i$ has the memoryless property.
(What does this imply?)
\end{itemize}
\end{frame}

\startframe{6.2 Continuous-Time Markov chains}
Examples:
\begin{itemize}
\item Any Poisson process
\item On-off process
\item Continuous-time random walk
\end{itemize}
\end{frame}

\startframe{6.2 Continuous-Time Markov chains}
Given a continuous-time, discrete-space MC, let us define
$P(t)$ to be the implied transition probability matrix at
time $t$.  What properties can we establish for $P(t)$?
(NB:  Not easy to compute $P(t)$ in general.)
\begin{itemize}
\item $P(0) = ??$
\item $\sum_j P_{ij}(t) = ??$
\item $P(s)P(t) = ??$
\end{itemize}
\end{frame}

\startframe{6.8 Computing the Transition Probabilities}
The generator (or rate) matrix $R$:
\begin{itemize}
\item Take $r_{ij} = \mbox{rate of $i\to j$ transitions}$, $i\ne j$.
\item Take $r_{ii} = -\sum_{j\ne i} r_{ij}$.
\item This implies $\sum_{j} r_{ij} = ???$
\end{itemize}
What is $R$ for the on-off process?
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


