\input{lecture}

\begin{document}


\def \thedate{Feb.~20}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item
For Wednesday,
read Sections 6.1 and 6.2.
\item 
HW\#5 is due {\em ELECTRONICALLY} on Friday.
\item 
See example code and .Rnw solutions files for homework
\item
Tentatively:  Midterm on Wednesday, Feb. 29 in evening.
\end{itemize}
\end{frame}

\startframe{5.3 The Poisson Process}
Proof of Poisson approximation to binomial:
\begin{itemize}
\item Poisson MGF:  $M(t) = \exp\{\lambda e^t - \lambda\}$
\item Binomial MGF:  $M(t) = (1+pe^t -p)^n$
\item Let $p=\lambda/n$ and see what happens as $n\to\infty$.
\end{itemize}
\end{frame}


\startframe{5.3 The Poisson Process}
\begin{itemize}
\item
Let $S_0=0$ and $S_1, S_2, \ldots$ are the times of events in a Poisson process.
\item
Let $T_i=S_i-S_{i-1}$ for $i=1, 2, \ldots$ be the times between events.
\item 
How are the $T_i$ distributed?
\end{itemize}
\end{frame}

\startframe{5.3 The Poisson Process}
\question{In a Poisson process with rate $\lambda$, what is the distribution of
the waiting time until the $n$th event?}
{Exponential}
{Gamma}
{Lognormal}
{Uniform}
\end{frame}

\startframe{5.3 The Poisson Process}
\question{Let $X_1, \ldots, X_n$ be independent exponential random variables
with rates $\lambda_1, \ldots, \lambda_n$, respectively, and $X_{(1)}=\min_{1\le i\le n} X_i$.
What is $P(X_1=X_{(1)})$?}
{$\lambda_1$}
{$\lambda_2+\cdots + \lambda_n$}
{$\lambda_1/(\lambda_1+\cdots + \lambda_n)$}
{$1/(\lambda_2+\cdots + \lambda_n)$}
\end{frame}

\startframe{5.3 The Poisson Process}
Poisson processes add:
\begin{itemize}
\item
If $N_1(t)$ and $N_2(t)$ are Poisson processes with rates $\lambda_1$ and $\lambda_2$, 
then $M(t) = N_1(t)+N_2(t)$ is a Poisson process with rate $\lambda_1+\lambda_2$.
\item
Similarly, suppose we label each event $M(t)$ as type-1 or type-2 with probabilities
$\lambda_1/(\lambda_1+\lambda_2)$ and 
$\lambda_2/(\lambda_1+\lambda_2)$, respectively.  Then the counting process for each
type of event is separately a Poisson process.
\end{itemize}
\end{frame}

\startframe{5.4.1 Nonhomogeneous Poisson Processes}
Notice a slight change from the homogeneous version:
\begin{itemize}
\item Counts of events in disjoint time intervals are independent.
\item $P[N(s+h)-N(s)=1] = \lambda(s) h + o(h)$ as $h\to 0$.
\item $P[N(s+h)-N(s)\ge2] = o(h)$ as $h\to 0$.
\item Technically, we should also say $N(0)=0$.
\end{itemize}
\end{frame}

\startframe{5.4.1 Nonhomogeneous Poisson Processes}
\begin{itemize}
\item
In a nonhomogeneous Poisson process,
$N(s+t)-N(s)$ is Poisson distributed with mean
\[
\int_s^{s+t} \lambda(y)\,dy.
\]
\item
If we define $m(t) = \int_0^t \lambda(y)\,dy$, then this is just
$m(s+t)-m(s)$.
\item
What is the distribution of the waiting time until the first event?
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


