\input{lecture}

\begin{document}


\def \thedate{Jan.~18}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
Read Section 4.1  before Friday's class.
\item
Office hours tomorrow will be 12:45 to 2:45 due to a comprehensive exam I have
to attend. 
\item There is a new dropbox that you can use to turn in homework, if you wish.
\end{itemize}
\end{frame}

\startframe{3.4 Computing Expectations by Conditioning}
\question{If $X\sim \mbox{Geom}(p)$, then $E(X) = $?}
{$\frac1p$}
{$\frac1{p^2}$}
{$\frac1{1-p}$}
{$\frac1{(1-p)^2}$}
\end{frame}

\startframe{3.4 Computing Expectations by Conditioning}
\begin{itemize}
\item Perform repeated independent Bernoulli$(p)$ experiments.
\item Let $N_k$ be the trial number where $k$ consecutive successes are first observed.
\item What is $E(N_k)$?
\end{itemize}
\end{frame}

\startframe{3.4 Computing Expectations by Conditioning}
\begin{itemize}
\item Suppose that adult males' heights have mean $\mu_M$ and variance
$\sigma^2_M$.
\item Similarly, females' heights have mean $\mu_F$ and variance $\sigma^2_F$.
\item If some proportion $p$ of the population is female, what is the variance of
the height of an adult selected at random from the population?
\end{itemize}
\end{frame}

\startframe{3.4 Computing Expectations by Conditioning}
\question{$\Var(X)$ equals which of the following?}
{$E \left[ E(X^2 \mid Y) \right]$}
{$E \left[ \Var(X \mid Y) \right]$}
{$\Var \left[ E(X \mid Y) \right]$}
{$E \left[ \Var(X \mid Y) \right] + \Var\left[ E(X\mid Y) \right]$}
\end{frame}

\startframe{3.5 Computing Probabilities by Conditioning}
\begin{columns}
\begin{column}{1.25in}
\tikz{
\def \lx{-1.4} \def \ux{2.4} \def \ly{-1.2} \def\uy{1.2}
%\draw (\lx,\ly) rectangle (\ux,\uy);
% grid with dot at origin for helping place objects in rectangle:
%\draw[step=.5, very thin] (\lx,\ly) grid (\ux,\uy); \filldraw (0,0) circle (.05);
%
%\draw (-1.2, 1) node {(c)};
\filldraw (-1,0) circle (.05);
\filldraw (2,0) circle (.05);
\draw (-1,.2) node {$A$};
\draw (2,.2) node {$B$};
\draw (-1,0) -- (-.75,0) -- (-.75, .75) -- (1.75,.75) -- (1.75,0) -- (2,0);
\draw (-1,0) -- (-.75,0) -- (-.75, -.75) -- (1.75,-.75) -- (1.75,0) -- (2,0);
\draw (.125,.75) -- (.125,0) -- (.875, 0) -- (.875, -.75);
%
\foreach \x / \y / \label  in {-.375 / .75 / 1, -.375 / -.75 / 2, 1.125 / .75 / 3, 1.125 / -.75 / 4, .375 / 0 / 5} 
  {\filldraw (\x, \y) circle (.05); \draw (\x, \y) -- (\x + .25,\y + .25); 
   \draw[white, thick] (\x + .05, \y) -- (\x + .25, \y); \draw (\x + .15, \y - .2) node {\small \label};
  }
}
\end{column}
\begin{column}{3.5in}
\begin{itemize}
\item
Switch $i$ is closed (i.e., current will flow) with probability $p_i$,
independently of others.
\item What is the probability that current will flow from $A$ to $B$?
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\startframe{3.5 Computing Probabilities by Conditioning}
\begin{itemize}
\item
At a party, $n$ men take off their hats, which are then mixed up.  
\item 
Each man randomly selects one of the hats.
\item 
What is the probability that nobody gets his own hat?
\end{itemize}
\end{frame}

\startframe{3.5 Computing Probabilities by Conditioning}
\begin{itemize}
\item 
Two players alternate flipping a coin that comes up heads with probability $p$.
\item
The first one to get heads is the winner.
\item 
What is the probability that the first player is the winner?
\end{itemize}
\end{frame}

%\startframe{1.2\quad Properties of Probability}
%
%\begin{columns}
%\begin{column}{4cm}
%
%\tikz{
%\def \lx{-1.4} \def \ux{2.4} \def \ly{-1.2} \def\uy{1.2}
%\draw (\lx,\ly) rectangle (\ux,\uy);
%% grid with dot at origin for helping place objects in rectangle:
%%\draw[step=.5, very thin] (\lx,\ly) grid (\ux,\uy); \filldraw (0,0) circle (.05);
%\draw (0,0) ellipse (1 and .8);
%\draw (1,0) ellipse (1 and .8);
%\draw (0,.9) node {\small$A$};
%\draw (1,.9) node {\small$B$};
%\draw (2.2,1) node{$S$};
%}
%\end{column}
%\begin{column}{6cm}
%Venn diagram depicting events $A$ and $B$.  In the picture, they appear to 
%have a nonempty intersection.
%\end{column}
%\end{columns}
%\end{frame}





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


