\input{lecture}

\begin{document}


\def \thedate{Feb.~3}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item
Please read Sections 5.1, 5.2.1, 5.2.2, and 5.2.3 (both editions) for Monday.
\item 
No class next Friday.
\end{itemize}
\end{frame}


\startframe{4.8 Time Reversible Markov Chains}
``Consider a stationary ergodic Markov chain (that is, an ergodic Markov chain
that has been in operation for a long time)\ldots''
\begin{itemize}
\item ergodic?
\end{itemize}
\end{frame}

\startframe{4.4 Limiting Probabilities}
Some terminology (with which you've already dealt to some extent!)
\begin{itemize}
\item period
\item aperiodic
\item positive recurrent
\item ergodic
\end{itemize}
\end{frame}

\startframe{4.3 Classification of States}
Remember this?

\question{If all of the states of a Markov chain communicate with each other,
the chain is said to be}
{irreducible}
{ergodic}
{recurrent}
{transient}
\end{frame}

\startframe{4.4 Limiting Probabilities}
Theorem:  For an irreducible ergodic (time-homogeneous) 
Markov chain with transition matrix $P$,
\[
\lim_{n\to\infty} P^n_{ij} 
\]
exists and it does not depend on $i$ (it does generally depend on $j$).
\end{frame}

\startframe{4.4 Limiting Probabilities}
Corollary:  For an irreducible ergodic (time-homogeneous) 
Markov chain with transition matrix $P$,
the equations 
\[
\pi_j = \sum_{i=0}^{\mbox{\# states}} \pi_i P_{ij}, \qquad j = 0, 1, 2, \ldots
\]
have solution 
\[
\pi_j=\lim_{n\to\infty} P^n_{ij}.
\]
\end{frame}

\startframe{4.4 Limiting Probabilities}
One last bit of notation:
\begin{itemize}
\item stationary probability
\end{itemize}
\end{frame}

\startframe{4.4 Limiting Probabilities}
More facts we will not prove:
\begin{itemize}
\item If we assume $\sum_j \pi_j=1$, then the solution in the corollary is unique.
\item The $\pi_j$ from the corollary are the stationary probabilities.
\item Aperiodicity is not necessary in order for the equations in the corollary to have a 
unique solution.  However, for irreducible chains, positive recurrence is necessary.
\item Stationary probabilities exist even for periodic irreducible chains.
\end{itemize}
\end{frame}

\startframe{4.8 Time Reversible Markov Chains}
\begin{itemize}
\item
Define $Q_{ij} = P(X_t = j \mid X_{t+1} = i)$.
\item
Use Bayes' theorem to prove 
\[
Q_{ij} = \frac{\pi_j P_{ji}}{\pi_i}.
\]
\item
To be continued\ldots
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


