\input{lecture}

\begin{document}


\def \thedate{Apr.~4}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
HW \#9 is due on Friday, April 6 at 2:30pm.
\item 
All homework must be turned in electronically from now on.
\item 
Something I had not noticed regarding Problem \#2:
You can use 5 uniforms to simulate a beta(3,2).  
\end{itemize}
\end{frame}


\startframe{Monte Carlo methods}
Importance sampling standard errors
\begin{itemize}
\item Plain importance sampling:  Easy
\item Ratio importance sampling:  Use delta method to derive
\[
\Var \left[ \sqrt{n} ( \tilde \mu - \mu)  \right] \approx \tilde\sigma^2,
\]
where $\tilde \sigma^2 = \alpha^2 (1,-\mu) \Sigma (1,-\mu)^\top$ and $\Sigma=\ldots$
\end{itemize}
\end{frame}


\startframe{Monte Carlo methods}
Importance Sampling Example:  What is $P(X>4.5)$ for $X\sim N(0,1)$?
\begin{itemize}
\item Try naive approach.
\item Try importance sampling with $q(x) = \exp\{-(x-4.5)\} I\{x>4.5\}$.
\item Try ratio importance sampling.  (Careful!)
\item This type of rare event problem often good for importance sampling.
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Lessons learned:
\begin{itemize}
\item Plain importance sampling needs %, $q$ must be chosen so that
$g(x)f(x)>0 \Rightarrow q(x)>0$.
\item Ratio importance sampling needs %, $q$ must be chosen so that
$f(x)>0 \Rightarrow q(x)>0$.
\item (Can use different samples for numerator 
and denominator.)
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Ratio importance sampling application:
\begin{itemize}
\item Goal:  Estimate $\mu(\theta) = \E_{f_\theta} g(X)$ for $\theta\in\Theta$.
\item Plain importance sampling:  Draw new sample for each $\theta\in\Theta$.
\item Ratio importance sampling:  Draw a {\em single} sample from $q(x)$\ldots

(Need: $f_\theta(x)>0$ implies $q(x)>0$ for all $\theta$.)
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Importance sampling can be unstable:
\begin{itemize}
\item Suppose $q(x)$ is small for some $x$ where $g(x)f(x)$ is big.  
\item Then $g(x)f(x)/q(x)$ will have extreme outliers that are rare.
\item If these values are rare enough, we may never see them and thus
we may underestimate the M.C. standard error.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}




