\input{lecture}

\begin{document}


\def \thedate{Mar.~16}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
For Monday, March 19:  Read Sections 6.6, 6.8
\item HW \#7 is due on Wednesday, March 21.  (Discuss?)
\end{itemize}
\end{frame}

\startframe{6.5 Limiting Probabilities}
Recall Section 4.4:
\begin{itemize}
\item
If limiting probabilities existed, so did unique stationary (long-term) probabilities.
\item Not conversely!  (Why not?)
\item In Chapter 6, life is simpler in a certain sense.
\end{itemize}
\end{frame}


\startframe{6.5 Limiting Probabilities}
\begin{itemize}
\item Suppose that $\lim_{t\to\infty}P_{ij}(t)$ exists and is independent of $i$ for all $i, j$.
\item Denote $P^\infty= \lim_{t\to\infty}P_{ij}(t)$ with each row the same.  (How to label?)
\item Derive:
$0 = P^\infty R$.
\end{itemize}
\end{frame}

\startframe{6.5 Limiting Probabilities}
The equations can be used to find the limiting $\pi$ vector:
\begin{itemize}
\item Another way to write $0 = P^\infty R$ is
\[
v_j\pi_j = \sum_{i\ne j} q_{ij}\pi_i
\quad\mbox{for all $j$.}
\]
\item Find $\pi$ if
\[
R=
\begin{bmatrix}
-\lambda_1 & \lambda_1 \\
\lambda_2 & -\lambda_2 
\end{bmatrix}.
\]
\end{itemize}
\end{frame}

\startframe{6.5 Limiting Probabilities}
Sufficient conditions for the existence of the limiting probabilities $\pi_j$:
\begin{itemize}
\item Irreducibility (all states communicate)
\item Positive recurrence (mean time to return to each state is finite)
\end{itemize}
When the limiting probabilities exist, we say the chain is ergodic.
\end{frame}

\startframe{6.5 Limiting Probabilities}
%\begin{itemize}
%\item
Re-interpretation of 
$v_j\pi_j = \sum_{i\ne j} q_{ij}\pi_i$
%\quad\mbox{
for all $j$:
\begin{itemize}
\item
In the long run,
the rate of leaving state $j$ is the same as the rate
of entering state $j$.
\item Apply this to birth-death processes.
\item Looking ahead to Section 6.6:  In fact, a birth-death process must be time-reversible.
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


