\input{lecture}

\begin{document}


\def \thedate{Feb.~15}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item
For Friday,
read the parts of Section 5.3 that you haven't yet finished.
\item 
Homework due this Friday; regular office hours Thursday.
\item
The midterm will be \ldots ???
\item 
The final will be Monday, Apr. 30 at 8:00am.
\end{itemize}
\end{frame}

\startframe{5.2 The exponential distribution}
Very important fact:  If $X_1$ and $X_2$ are independent 
exponential random variables with rates $\lambda_1$ and
$\lambda_2$, then
\[
P(X_1 < X_2) = \frac{\lambda_1}{\lambda_1 + \lambda_2}
\]
{\em Where did we use the independence?}
\end{frame}

\startframe{5.2 The exponential distribution}
\question{The minimum of a set of $n$ i.i.d.~exponential random variables has what type of distribution?}
{Exponential}
{Lognormal}
{Normal}
{Gamma}
\end{frame}

\startframe{5.2 The exponential distribution}
Suppose you arrive at a post office having two clerks at 
a moment when both are busy but nobody else is waiting in line.  
The service times for clerk $i$ are exponential with rate $\lambda_i$.
What is the expected amount of time you have to wait in line?
\end{frame}

\startframe{5.3 The Poisson Process}
Counting processes:
\begin{itemize}
\item $N(t)=$number of ``events'' that occur in $[0, t]$.
\item $N(t)$ is integer-valued and nondecreasing in $t$.
\item Consider an example\ldots
\end{itemize}
\end{frame}

\startframe{5.3 The Poisson Process}
A Poisson process is a special type of counting process:
\begin{itemize}
\item Counts of events in disjoint time intervals are independent.
\item For any $0\le s<t$, the number of events in $(s,t]$ is Poisson with parameter
$\lambda(t-s)$.
\item Technically, we should also say $N(0)=0$.
\end{itemize}
\end{frame}

\startframe{5.3 The Poisson Process}
A Poisson process is a special type of counting process
(equivalent re-definition):
\begin{itemize}
\item Counts of events in disjoint time intervals are independent.
\item $P[N(s+h)-N(s)=1] = \lambda h + o(h)$ as $h\to 0$.
\item $P[N(s+h)-N(s)\ge2] = o(h)$ as $h\to 0$.
\item Technically, we should also say $N(0)=0$.
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


