\input{lecture}

\begin{document}


\def \thedate{Apr.~2}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
HW \#9 is due on Friday, April 6 at 2:30pm.
\item 
All homework must be turned in electronically from now on.
\item Problem \#2 is not as straightforward as I had intended, but go ahead and solve it as asked.
This may involve numerically finding the root of a 4th-degree polynomial.
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Ratio Importance Sampling:
\begin{itemize}
\item Goal:  Find $\mu = \int g(x)f(x)\,dx = \E_f g(X)$.
\item Ratio estimator is
\[
\tilde \mu = 
\frac{\frac1n \sum_i  g(X_i)r(X_i)/q(X_i)}{\frac1n \sum_i  r(X_i)/q(X_i)}
\]
where $X_1, \ldots, X_n \stackrel{\rm iid}{\sim} q$ and $r(x) = f(x)/\alpha$.
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Ratio importance sampling.
\begin{itemize}
\item Unlike $\hat\mu=\frac1n \sum_i  g(X_i)f(X_i)/q(X_i)$, $\tilde \mu$
is not unbiased. 
\item However, $\mbox{MSE}(\tilde \mu) \le \mbox{MSE}(\hat\mu)$.

( J. Liu, {\it Monte Carlo Strategies in Scientific Computing})
\item Must assume that $f(x)>0$ implies $q(x)>0$.
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Importance sampling standard errors
\begin{itemize}
\item Plain importance sampling:  Easy
\item Ratio importance sampling:  Use delta method to derive
\[
\Var \left[ \sqrt{n} ( \tilde \mu - \mu)  \right] \approx \tilde\sigma^2,
\]
where $\tilde \sigma^2 = (1,-\mu) \Sigma (1,-\mu)^\top$ and $\Sigma=\ldots$
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Importance Sampling Example:  What is $P(X>4.5)$ for $X\sim N(0,1)$?
\begin{itemize}
\item Try naive approach.
\item Try importance sampling with $q(x) = \exp\{-(x-4.5)\} I\{x>4.5\}$.
\item Try ratio importance sampling.  (Careful!)
\end{itemize}
\end{frame}

\startframe{Monte Carlo methods}
Ratio importance sampling application:
\begin{itemize}
\item Goal:  Estimate $\mu(\theta) = \E_{f_\theta} g(X)$ for $\theta\in\Theta$.
\item Plain importance sampling:  Draw new sample for each $\theta\in\Theta$.
\item Ratio importance sampling:  Draw a single sample from $q(x)$\ldots

(Need: $f_\theta(x)>0$ implies $q(x)>0$ for all $\theta$.)
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


