\input{summary}

\begin{document}


\def \thedate{Feb.~3}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item
Please read Sections 5.1, 5.2.1, 5.2.2, and 5.2.3 (both editions) for Monday.
\item 
No class next Friday.
\item Office hours will be moved from Thursday, Feb. 9 to Tuesday, Feb. 7 from 2:00 to 4:00.
\end{itemize}
\notes{We will not begin Chapter 5 on Monday; that will most likely happen on Wednesday.}
\end{frame}


\startframe{4.8 Time Reversible Markov Chains}
``Consider a stationary ergodic Markov chain (that is, an ergodic Markov chain
that has been in operation for a long time)\ldots''
\begin{itemize}
\item ergodic?
\end{itemize}
\notes{This is the very first line from Section 4.8, and so it's time to talk about what 
``ergodic'' means.}
\end{frame}

\startframe{4.4 Limiting Probabilities}
Some terminology (with which you've already dealt to some extent!)
\begin{itemize}
\item period
\item aperiodic
\item positive recurrent
\item ergodic
\end{itemize}
\notes{Each of the last three of these is a class property.  }
\end{frame}

\startframe{4.3 Classification of States}
Remember this?

\question{If all of the states of a Markov chain communicate with each other,
the chain is said to be}
{irreducible}
{ergodic}
{recurrent}
{transient}
\notes{Correct answer:  A.  }
\end{frame}

\startframe{4.4 Limiting Probabilities}
Theorem:  For an irreducible ergodic (time-homogeneous) 
Markov chain with transition matrix $P$,
\[
\lim_{n\to\infty} P^n_{ij} 
\]
exists and it does not depend on $i$ (it does generally depend on $j$).
\notes{This is a very important theorem!}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


