\input{summary}

\begin{document}


\def \thedate{Feb.~29}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
Midterm exam:  7:00pm tonight, Feb.~29 in 105 Willard.
(NB:  Change of room!)
\item 
HW \#6 will be due on Friday (Mar.~2) at 2:30.
\end{itemize}
\notes{I discussed the beta distribution as it relates to exercise \#1
on your homework.  This discussion is summarized in the example code file
that you will find if you go to the course homework page.
}
\end{frame}

\startframe{6.2 Continuous-Time Markov chains}
Given a continuous-time, discrete-space MC, let us define
$P(t)$ to be the implied transition probability matrix at
time $t$.  What properties can we establish for $P(t)$?
(NB:  Not easy to compute $P(t)$ in general.)
\begin{itemize}
\item $P(0) = ??$
\item $\sum_j P_{ij}(t) = ??$
\item $P(s)P(t) = ??$
\end{itemize}
\notes{Using the definition $P_{ij}(t) = P[X(t)=j \mid X(0)=i]$,
we argued that $P(0)$ is the identity, each row of $P(t)$ sums
to one, and $P(s)P(t)=P(s+t)$.  That last fact is due to the Kolmogorov-Chapman
equations.}
\end{frame}

\startframe{6.8 Computing the Transition Probabilities}
The generator (or rate) matrix $R$:
\begin{itemize}
\item Take $r_{ij} = \mbox{rate of $i\to j$ transitions}$, $i\ne j$.
\item Take $r_{ii} = -\sum_{j\ne i} r_{ij}$.
\item This implies $\sum_{j} r_{ij} = \ldots$
\end{itemize}
What is $R$ for the on-off process?  A homogeneous Poisson process?
\notes{We discussed the equation $P(s)P(t)=P(s+t)$ and how it suggests
that $P(t)$ is some kind of a ``matrix exponential'' function.  In a certain sense,
we will see that $P(t)$ is $\exp\{Rt\}$.
}
\end{frame}

\startframe{6.4 The Transition Probability Function}
The entries of $R$ were introduced much earlier (but not in this form).
\begin{itemize}
\item Take $q_{ij} = \mbox{rate of $i\to j$ transitions}$, $i\ne j$.
\item Take $v_{i}$ so that $P_{ij}=q_{ij}/v_{i}$.
\item Clearly this implies $v_i=\sum_{j\ne i} q_{ij}$.
\end{itemize}
\notes{In the book, $q_{ij}$ is used for $r_{ij}$ when $j\ne i$
and $v_i=-r_{ii}$.}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


