\input{summary}

\begin{document}


\def \thedate{Mar.~16}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
For Monday, March 19:  Read Sections 6.6, 6.8
\item HW \#7 is due on Wednesday, March 21.  (Discuss?)
\end{itemize}
\notes{We decided that the due date will be Friday, March 23 instead 
because it seems to be the preference of most students and 
because I have no reason not to do this.  I have made the change on the course web site.}
\end{frame}

\startframe{6.5 Limiting Probabilities}
Recall Section 4.4:
\begin{itemize}
\item
If limiting probabilities existed, so did unique stationary (long-term) probabilities.
\item Not conversely!  (Why not?)
\item In Chapter 6, life is simpler in a certain sense.
\end{itemize}
\notes{
In the discrete-time case, a Markov chain could be periodic, which destroys the possibility
of limiting probabilities even though a unique set of stationary probabilities can exist.
In the continuous-time case, however, periodicity cannot occur, so this issue does not
arise.
}
\end{frame}


\startframe{6.5 Limiting Probabilities}
\begin{itemize}
\item Suppose that $\lim_{t\to\infty}P_{ij}(t)$ exists and is independent of $i$ for all $i, j$.
\item Denote $P^\infty= \lim_{t\to\infty}P_{ij}(t)$ with each row the same.  (How to label?)
\item Derive:
$0 = P^\infty R$.
\end{itemize}
\notes{
The derivation of $0 = P^\infty R$ begins with the forward Kolmogorov equations.
We discussed the interchange of the limit ($t\to\infty$) and the summation 
(in $P^\infty R$); I mentioned that this interchange could prove problematic
if we started with the backward Komogorov equations.  However, there is another,
more obvious,  reason that the  backward equations don't lead to a useful
equation here.  I overlooked this!  But we'll discuss it in Monday's class.
}
\end{frame}

\startframe{6.5 Limiting Probabilities}
The equations can be used to find the limiting $\pi$ vector:
\begin{itemize}
\item Another way to write $0 = P^\infty R$ is
\[
v_j\pi_j = \sum_{i\ne j} q_{ij}\pi_i
\quad\mbox{for all $j$.}
\]
\item Find $\pi$ if
\[
R=
\begin{bmatrix}
-\lambda_1 & \lambda_1 \\
\lambda_2 & -\lambda_2 
\end{bmatrix}.
\]
\end{itemize}
\notes{
We did not discuss the first equation---this will happen on Monday---but the second one 
(the rate matrix for the on-off process)
leads to 
\[
\pi_1=\frac{\lambda_2}{\lambda_1+\lambda_2} \quad\mbox{and}\quad
\pi_2=\frac{\lambda_1}{\lambda_1+\lambda_2}.
\]
}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}


