\input{summary}

\begin{document}


\def \thedate{Apr.~27}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\startframe{Announcements}
\begin{itemize}
\item 
Make sure you do not communicate with anyone (except me) 
about the take-home final.
\item
The in-class final will be held in 075 Willard (Not our normal classroom!)
\end{itemize}
\notes{We spent the whole day discussing the outline of the Monte Carlo 
and MCMC sections of the course.  The outline appears on the third page
of these notes.  On the second page is a broader outline of the
entire course.}
\end{frame}


\startframe{STAT 515 brief recap}
\begin{itemize}
\item Chapters 1 and 2:  Review of probability concepts
\item Chapter 3:  Conditional probability and conditioning (omit 3.6, 3.7)
\item Chapter 4:  Disc.-time, disc.-space Markov chains (omit 4.10, 4.11)
\item Chapter 5:  Exp. distribution, Pois. processes (omit some of 5.3, 5.4)
\item Midterm exam:  Chapters 1--5
\item Chapter 6:  Cont.-time, disc.-space Markov chains (omit 6.7)
\item Monte Carlo methods
\item Markov chain Monte Carlo
\end{itemize}
\notes{
Roughly speaking,  you should expect 5 to 10 points (out of 30) on the final exam to cover
the first part of the course (before the midterm).  You should expect the remaining 20 to 25
points to cover the second (post-midterm) part of the course.
}
\end{frame}

\startframe{STAT 515 brief recap}
More detail on the last two\ldots
\begin{itemize}
\item Monte Carlo methods
  \begin{itemize}
  \item Understanding what ``pseudo-random'' numbers are
  \item Inversion method
  \item Rejection method
  \item Simulating multivariate normals
  \item Importance Sampling
  \item Ratio importance sampling
  \item Standard error estimation
  \end{itemize}
\item Markov chain Monte Carlo
  \begin{itemize}
  \item Metropolis-Hastings algorithms
    \begin{itemize}
    \item Special case:  Symmetric proposal (Metropolis alg.)
    \item Special case:  Propose from full cond. (Gibbs samp.)
    \end{itemize}
  \item Disc.-time, cont.-space Markov chains
    \begin{itemize}
    \item Markov transition density $k ( y\mid x)$
    \item Stationary density $\pi(x)$.
    \item Harris ergodicity ($\pi$-irreducibility, aperiodicity, Harris recurrence)
    \end{itemize}
  \item Variable-at-a-time MH; full conditionals
  \item MCMC standard errors using batch means
  \item Effective sample size
  \end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}




